Estimating the Initial Susceptible Fraction in Dengue Dynamics

### 2014-10-08 draft_lm1.tex
minor changes

To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
Carlo techiniques implemented in the Bayesian inference with Python 
(BIP)~\cite{pone2011} available at 
\url{http://code.google.com/p/bayesian-inference/}.
BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
scheme that efficiently samples from high-dimensional joint distributions using 
multiple adaptive chains running in parallel with delayed rejection.
In this study we used one chain per parameter, i.e, 3 chains for each run.
The chains were run until $5000$ samples were obtained after discarding $500$ 
burn-in samples.
Convergence of the parallel chains was verified at every 100 iterations by the 
calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
% Results and Discussion can be combined.
\section*{Results}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
%\subsection*{Subsection 1}
\section*{Discussion}
% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
posterior inference for $R(t)$.

>>>>

To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
Carlo techiniques implemented in the Bayesian inference with Python 
(BIP)~\cite{pone2011} available at 
\url{http://code.google.com/p/bayesian-inference/}.
BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
scheme that efficiently samples from high-dimensional joint distributions using 
multiple adaptive chains running in parallel with delayed rejection.
In this study we used one chain per parameter, i.e, 3 chains for each run.
The chains were run until $5000$ samples were obtained after discarding $500$ 
burn-in samples.
Convergence of the parallel chains was verified at every 100 iterations by the 
calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
% Results and Discussion can be combined.
\section*{Results}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
%\subsection*{Subsection 1}
\subsection{Serotype proportion}
The susceptible population estimated in this paper is defined as individuals 
susceptible to at least one of the circulating serotypes in any given year. 
This is due to the fact that we use case data to estimate $S_0$ prior to the 
epidemic. In general, susceptibility to dengue may be represented by 
4 binomially distributed random variables. Each representing the 
fraction of susceptibles to each virus type -- $S_i \sim Bin(p_i), \, i \in  
\{1, 2, 3, 4\}$. In the absence of 
cross-immunity, the suceptibility can be manifested, in individuals, as one of 
the following forms:
\begin{equation}
 S = \{S_{1}, S_{2}, S_{3}, 
S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
\end{equation}

Even though we don't observe the incidences for each serotype independently, we 
have information about the proportion of Serotypes, based on the lab confirmed 
cases. We also know that the drop in susceptibles estimated by the model, 
corresponds to the drop in the number of the one or more serotypes. For 
example: in an epidemic of just DENV1, the debt of susceptibles would have to 
be distributed between the following types: 
$\{S_{1},S_{12},S_{123},S_{1234},S_{13},S_{14}\}$.

\section*{Discussion}
% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
posterior inference for $R(t)$.
